What dictates the quality of Red Wine by Matthew Morris
========================================================

In order to understand the wine variables, I will plot distribution of each 
variable. This will give me an idea if there is normal distrubution, bi-modal, 
skews or outliers that will need to be addressed. 

References
Wine Understanding

Quality wine lables 
https://www.winespectator.com/articles/scoring-scale

Acidity in wine ( fixed.acidity, volatile.acidity,ph) 
https://winefolly.com/deep-dive/understanding-acidity-in-wine/

Sugar levels and wine
http://www.naijawinelovers.com/what-is-residual-sugar-in-wine/

Alcohol levels and wine
https://winefolly.com/tips/the-lightest-to-the-strongest-wine/

pH
https://winemakermag.com/technique/1650-monitoring-adjusting-ph


R references

https://mgimond.github.io/ES218/Week09a.html
https://rpubs.com/Mentors_Ubiqum/scale_x_continuous
http://www.sthda.com/english/wiki/ggally-r-package-extension-to-ggplot2-for-correlation-matrix-and-survival-plots-r-software-and-data-visualization
http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
http://www.sthda.com/english/wiki/qplot-quick-plot-with-ggplot2-r-software-and-data-visualization
https://stackoverflow.com/questions/38446804/coloring-points-based-on-variable-with-r-ggpairs
https://briatte.github.io/ggcorr/


# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
#Load all the required libraries
library("ggplot2")
library("dplyr")
library("gridExtra")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Loading the csv file and summary
redwine <- read.csv('redwinedata.csv')
```

# Initial impressions of distribution

```{r echo=FALSE, message=FALSE, warning=FALSE}

# fixed.acidity

grid.arrange(
qplot(x = fixed.acidity, data = redwine) +
  geom_histogram(binwidth = .25) +
  scale_x_continuous(breaks=.25:16),
# Violin plot
qplot(X, fixed.acidity, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)

```

##discovery - majority of fixed acidity population between 5.75 and 10.75

```{r echo=FALSE, message=FALSE, warning=FALSE}

# volatile.acidity
grid.arrange(
qplot(x = volatile.acidity, data = redwine) +
  geom_histogram(binwidth = .05) +
  scale_x_continuous(breaks = seq(from = .01, to = 1.1, by = .25)),
qplot(X, volatile.acidity, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)
```

## discovery - majority of volatile population between .25 and .75 

```{r echo=FALSE, message=FALSE, warning=FALSE}

# citric.acid
grid.arrange(
qplot(x = citric.acid, data = redwine) +
  geom_histogram(binwidth = .01) +
  scale_x_continuous(breaks = seq(from = .0, to = 1, by = .1)),
qplot(X, citric.acid, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)
```

## discovery - citric acid has a slight bimodal distribution between .0 and .25 and .35 and .5

```{r echo=FALSE, message=FALSE, warning=FALSE}

# residual.sugar

grid.arrange(
qplot(x = residual.sugar, data = redwine) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(from = .9, to = 15.7, by = 2)),
qplot(X, residual.sugar, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)


```

## discovery - residual.sugar is skewed right with outliers population between 1 and 3 

```{r echo=FALSE, message=FALSE, warning=FALSE}

#Creating a new Factored Variable called 'sugar_rating'

redwine$sugar_rating <- factor(redwine$residual.sugar)


redwine$sugar_rating <- ifelse(redwine$residual.sugar <=.3,'Bone Dry',
                      ifelse(redwine$residual.sugar <= 1.7,'Dry',
                      ifelse(redwine$residual.sugar <= 3.5,'Off Dry',
                      ifelse(redwine$residual.sugar <= 12,'Med Sweet',
                      ifelse(redwine$residual.sugar > 12, 'Sweet', 'no rating'
                             )))))
redwine$sugar_rating <- ordered(redwine$sugar_rating,
                       levels = c('Bone Dry', 'Dry', 'Off Dry',
                                  'Med Sweet', 'Sweet', 
                                  'no rating'))


ggplot(data = redwine, aes(x=sugar_rating))+
  geom_bar()

```

## Data set is dominated by Off Dry wines


```{r echo=FALSE, message=FALSE, warning=FALSE}    

# chlorides
grid.arrange(
qplot(x = chlorides, data = redwine) +
  geom_histogram(binwidth = .01) +
  scale_x_continuous(breaks = seq(from = .01, to = .7, by = .1)),
qplot(X, chlorides, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)
```

## discovery - chlorides are skewed right with outliers 

```{r echo=FALSE, message=FALSE, warning=FALSE}  

# free.sulfer.dioxide
grid.arrange(
qplot(x = free.sulfur.dioxide, data = redwine) +
  geom_histogram(binwidth = 10) +
  scale_x_continuous(breaks = seq(from = 10, to = 72, by = 10)),
qplot(X, free.sulfur.dioxide, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)

```

## discovery - skewed right with outliers

```{r echo=FALSE, message=FALSE, warning=FALSE}

#total.sulfer.dioxide
grid.arrange(
qplot(x = total.sulfur.dioxide, data = redwine) +
  geom_histogram(binwidth = 40) +
  scale_x_continuous(breaks = seq(from = 40, to = 290, by = 40)),
qplot(X, total.sulfur.dioxide, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)
  
```

## discovery - total.sulfer.dioxide skewed right with outliers 

```{r echo=FALSE, message=FALSE, warning=FALSE}  

# density
grid.arrange(
qplot(x = density , data = redwine) +
  geom_histogram(binwidth = 0.001) +
  scale_x_continuous(breaks = seq(from = .01, to = 2, by = .25)),
qplot(X, density, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)

```  

## discovery - fairly normal distribution for density

```{r echo=FALSE, message=FALSE, warning=FALSE}  

# pH
grid.arrange(
qplot(x = pH , data = redwine) +
  geom_histogram(binwidth = .15) +
  scale_x_continuous(breaks = seq(from = 2.73, to = 4, by = .25)),
qplot(X, pH, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)
```

## discovery - pH has a normal distribution

```{r echo=FALSE, message=FALSE, warning=FALSE}  

# sulphates
grid.arrange(
qplot(x = sulphates , data = redwine) +
  geom_histogram(binwidth = .1) +
  scale_x_continuous(breaks = seq(from = .33, to = 2, by = .5)),
qplot(X, sulphates, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)
```  

## discovery - sulphates skewed right with outliers 
## fairly normal distribution between .33 and .93

```{r echo=FALSE, message=FALSE, warning=FALSE}  

# alcohol
grid.arrange(
qplot(x = alcohol , data = redwine) +
  geom_histogram(binwidth = .5) +
  scale_x_continuous(breaks = seq(from = 8.4, to = 14.9, by = 1)),
qplot(X, alcohol, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)

```  

## discovery - alcohol us skewed right with outliers 
## potentially outliers to right and left

```{r echo=FALSE, message=FALSE, warning=FALSE}  

redwine$alcohol_rating <- factor(redwine$alcohol)

redwine$alcohol_rating <- ifelse(redwine$alcohol <10,'Low',
                      ifelse(redwine$alcohol < 11.5,'Med Low',
                      ifelse(redwine$alcohol < 13.5,'Medium',
                      ifelse(redwine$alcohol < 15,'Med High',
                      ifelse(redwine$alcohol >= 15, 'High', 'no rating'
                             )))))
redwine$alcohol_rating <- ordered(redwine$alcohol_rating,
                       levels = c('Low', 'Med Low', 'Medium',
                                  'Med High', 'High', 'no rating'))

ggplot(data = redwine, aes(x=alcohol_rating))+
  geom_bar()


```

## The data set is primarly low to med low alcohol rating

```{r echo=FALSE, message=FALSE, warning=FALSE}  

#quality
grid.arrange(
qplot(x = quality , data = redwine) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(from = 3, to = 8, by = 1)),
qplot(X, quality, data = redwine, 
      geom=c("violin"), trim = FALSE), nrow =1)
 
```

## discovery - quality has fairly normal distribution
I will need to create these levels as factors as this will be my category to 
compare features to. 

```{r echo=FALSE, message=FALSE, warning=FALSE}  
redwine$quality_rating <- factor(redwine$quality)


redwine$quality_rating <- ifelse(redwine$quality ==3, 'Not recommended', 
               ifelse(redwine$quality ==4, 'Mediocre',
               ifelse(redwine$quality ==5, 'Good',
               ifelse(redwine$quality ==6, 'Very good',       
               ifelse(redwine$quality ==7, 'Outstanding',        
               ifelse(redwine$quality ==8, 'Classic', 'no rating'
               ))))))
redwine$quality_rating <- ordered(redwine$quality_rating,
                       levels = c('Not recommended', 'Mediocre', 'Good',
                                  'Very good', 'Outstanding', 'Classic', 
                                  'no rating'))
ggplot(data = redwine, aes(x=quality_rating))+
  geom_bar()

```

## The data set is mostly Good and Very Good Wines

# Univariate Analysis

### What is the structure of your dataset?
After reviewing all of the variables it became apparent that it would quality 
may be predicted by the other variables in the data set.

After doing some research online I found that navigating the relationships of 
acidity, Chlorides, sulfates and pH is fairly complex balancing act. 
Most of these factors are right skew and while I would be tempted to remove 
outliers from this I feel I would need to research more on how all of the 
variations of acidity, chlorides, sulfates and pH play a factor in quality. 
While I may be able to find positive or negative corrleations I would not feel
confident in reporting on those knowing that other factors may balances those
out ie lurking variables. 

### What is/are the main feature(s) of interest in your dataset?
Because of the complexity mentioned above, my initial focus
will be on Sugars and alcohol and how they play a part in Quality rating. 

Essentially does sweeter or more boozier wines tend to have a higher quality
in this data set.
Becuase most wines are limited to Good and Very Good this could pose 
a challenge.


### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

As I have time I would like to explore more of the chemistry and associations 
between acidity, chlorides, sulfates and pH. I have dones some research online
on density but have not found much useful at this stage. 

### Did you create any new variables from existing variables in the dataset?
I created  3 new factor variables to label; quality, residual.sugar, 
and alcohol. These are; quality_rating, sugar_rating and alcohol_rating. 
I used cited reference to create the rating system. 

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
Holestically looking at the distribution if they are skewed it is always to the
right. 
Factors created show major dominance in just 1 to 2 areas rather than spread
throughout. 
Sugar.residual == Off Dry
alcohol == Low, Med Low
Quality == Good, Very Good
This tells me we are mostly looking at 1 or 2 varietals in the data set.
With enough time a model could be created wherein you could predict what 
varietal the wine is based on the variables in this data set. 

##### volatile.acidity: 
Distribution: somewhat normal, skew right 
Majority population: between .25 and .75

#### citric.acid 
Distribution: bimodal distribution, slight skew right
Majority of population: Between .0 and .25 and .35 and .5
#### residual.sugar
Distribution: Skewed Right
Majority of population: Between 1 and 3

##### chlorides
Distribution: skewed right with outliers 
Majority of population: need more exploration detail
#### free.sulfer.dioxide
Distribution: Skewed Right
Majority of population: Between 5 and 35

#### total.sulfer.dioxide
Distribution: Skewed Right
Majority of population: Between 40 and 80

#### Density
Distribution: Normal
Majority of population: Between .994 and 1.0

#### pH
Distribution: Normal
Majority of population: Between 3.03 and 3.63
#### Sulphates
Distribution: Skewed Right
Majority of population: Between .33 and .93

#### Alcohol 
Distribution: Skewed Right
Majority of population: Between 8.8 and 11.9

#### Quality
Distribution: categorical and focused primarily on 2
Majority of population: Category 5, 6





# Bivariate Plots Section
It would be interesting to see how sugar and alcohol relate to quality. I would
also like to see how sugar relates to alcohol. 

# Residual.sugar to Quality_rating
```{r echo=FALSE, message=FALSE, warning=FALSE}  

grid.arrange(
ggplot(redwine, aes(x = quality_rating, y = residual.sugar)) + geom_point() + 
  stat_smooth(method = "lm", se = FALSE, formula = y ~  x + I(x^2) ),
  
  
ggplot(data = redwine, aes(x = quality_rating, y = residual.sugar)) +
  geom_jitter( alpha = .2) +
  geom_boxplot(alpha = .3,color = 'blue') +
  stat_summary(fun = "mean", 
               geom = "point", 
               color = "red", 
               shape = 1, 
               size = 4), nrow =1)
```

## Closer look into relationships between quality and alcohol.

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library("GGally")
rwdf <-data.frame(redwine$residual.sugar, redwine$quality_rating)
ggpairs(rwdf)
```

Good and Very Good Wines tend to have more sugar in them as opposed to;
Not recommended, Mediocre, Outstanding and Classic. This still could be due to 
lack of observations in the data set. It does appear that there is some 
relationship between more sugar and quality.


```{r echo=FALSE, message=FALSE, warning=FALSE} 

grid.arrange(
ggplot(redwine, aes(x = quality_rating, y = alcohol)) + geom_point() + 
  stat_smooth(method = "lm", se = FALSE, formula = y ~  x + I(x^2) ),
  
  
ggplot(data = redwine, aes(x = quality_rating, y = alcohol)) +
  geom_jitter( alpha = .2) +
  geom_boxplot(alpha = .3,color = 'blue') +
  stat_summary(fun = "mean", 
               geom = "point", 
               color = "red", 
               shape = 1, 
               size = 4), nrow =1)
  
```

## Closer look into relationships between quality and alcohol.

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library("GGally")
rwdf <-data.frame(redwine$alcohol, redwine$quality_rating)
ggpairs(rwdf)
```
There appears to be a relationship between alcohol and quality of wine. The data
here also looks to be more evenly distriubuted than residual.sugar.

```{r echo=FALSE, message=FALSE, warning=FALSE} 

grid.arrange(
ggplot(redwine, aes(x = alcohol, y = residual.sugar)) + geom_point() + 
  stat_smooth(method = "lm", se = FALSE, formula = y ~  x + I(x^2) ),
  
  
ggplot(data = redwine, aes(x = alcohol, y = residual.sugar)) +
  geom_jitter( alpha = .2) +
  geom_boxplot(alpha = .3,color = 'blue') +
  stat_summary(fun = "mean", 
               geom = "point", 
               color = "red", 
               shape = 1, 
               size = 4), nrow =1)
```

## Closer look into relationships between residual.sugar, alcohol, and quality.

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library("GGally")
rwdf <-data.frame(redwine$residual.sugar, redwine$alcohol)
ggpairs(rwdf)
```
In this dataset there is more sugars in lower alcohol wines. Supporting that 
alcohol may indeed play a factor in quality.

## Other features review

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library("GGally")
ggcorr(data = redwine, title = "Red Red Wine")
```
At first review it would look as though alcohol has the strongest relationship
to quality. There are other features that have significant relationships that I believe are worth reviewing. While I still need a deeper business understanding of acidity, chlorides, sulfates, pH, and density I believe it is worth
at least reviewing for now and compare to alcohols relationship to quality. 

```{r echo=FALSE, message=FALSE, warning=FALSE} 
cor_test <- function(x, y) {
  return(cor.test(x, as.numeric(y))$estimate)
}
correlations <- c(
  cor_test(redwine$fixed.acidity, redwine$quality_rating),
  cor_test(redwine$volatile.acidity, redwine$quality_rating),
  cor_test(redwine$citric.acid, redwine$quality_rating),
  cor_test(log10(redwine$residual.sugar), redwine$quality_rating),
  cor_test(log10(redwine$chlorides), redwine$quality_rating),
  cor_test(redwine$free.sulfur.dioxide, redwine$quality_rating),
  cor_test(redwine$total.sulfur.dioxide, redwine$quality_rating),
  cor_test(redwine$density, redwine$quality_rating),
  cor_test(redwine$pH, redwine$quality_rating),
  cor_test(log10(redwine$sulphates), redwine$quality_rating),
  cor_test(redwine$alcohol, redwine$quality_rating))
names(correlations) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',
                         'log10.residual.sugar',
                         'log10.chlordies', 'free.sulfur.dioxide',
                         'total.sulfur.dioxide', 'density', 'pH',
                         'log10.sulphates', 'alcohol')
correlations
```

Alcohol still seems to have the highest relationship on quality, 
followed by volatile.acidity. 
# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
My features of interest were residual, sugar, alcohol and quality_rating. 
My analysis shows that there is a lack of evenly distributed variety of 
observations which I believe is hindering some of the analysis. As for now, 
Alchol has a stronger correlation to quality than sugar and while there may be
some relationship to residual.sugar and alcohol it is low with a correlation 
score of .042



### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Relationship features of interest or significance are defined here as 
having a correlation score of -+ .63 and higher/lower respective to stronger r
scores.
summary(redwine)
fixed.acidity to citric.acid rscore .67
fixed.acidity to density rscore .67
fixed.acidity to pH -.68

The top 3 correlations to Quality are
alcohol to quality rscore .48
volatile.acidity to quality rscore .39
sulphate to quality rscore .25

feature of note sugar, I was reviweing this and found the 
correlation is not significant wtih an rscore of .014

I would expect acidity to have a strong relationship with pH. I expect there is more to be learned about how, chlorides,
acidity, sulphates and density work together to build or lower quality. I'm still theorizing that within even 1 variatal sugar and alcohol together play a strong role in quality. 

### What was the strongest relationship you found?
The strongest relationship in the dataset is pH. to fixed acidity. The strongest relationship to quality is alcohol



# Multivariate Plots Section

> Knowing that Alcohol is a major contributer to overall quality, I would like
to see what combines with alcohol building a stronger relationship to quality.
My initial hunch is we will see a relationship with citric.acidity and or sugar. 


```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine,
       aes(y = fixed.acidity, x = alcohol,
           color = quality_rating)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~quality_rating) +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality'))
```

Fixed.acidity and alcohol has some slight positive coorelations in Good wines.
Overall the relationship is negative. 

```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine,
       aes(y = volatile.acidity, x = alcohol,
           color = quality_rating)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~quality_rating) +
  scale_color_brewer(type='seq',
      guide=guide_legend(title='Quality'))
```
While volatile.acidity and alcohol do not have much of a correlation in quality
of wines, there looks to be a positive correlation in Classic wines. 

```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine,
       aes(y = citric.acid, x = alcohol,
           color = quality_rating)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~quality_rating) +
  scale_color_brewer(type='seq',
      guide=guide_legend(title='Quality'))
```
Citric.acid and alcohol has some slight positive coorelations in 
Very Good wines.
Overall the relationship is negative. There is a strong negative Correlations 
With classic wines.

```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine,
       aes(y = residual.sugar, x = alcohol,
           color = quality_rating)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~quality_rating) +
  scale_color_brewer(type='seq',
      guide=guide_legend(title='Quality'))
```


While there is a slight positive correlation with residual.sugar and Alcohol
with Quality of wine it is strongest with Mediocre and Not recommended wines.

```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine,
       aes(y = chlorides, x = alcohol,
           color = quality_rating)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~quality_rating) +
  scale_color_brewer(type='seq',
      guide=guide_legend(title='Quality'))
```
Chlorides and Alcohol overall have a negative correlation with quality of wine. 


```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine,
       aes(y = free.sulfur.dioxide, x = alcohol,
           color = quality_rating)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~quality_rating) +
  scale_color_brewer(type='seq',
      guide=guide_legend(title='Quality'))
```

Here again, free.sulfur.dioxide and alcohol have overal negative correlations
and then a positve one in classic quality rating. 

```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine,
       aes(y = total.sulfur.dioxide, x = alcohol,
           color = quality_rating)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~quality_rating) +
  scale_color_brewer(type='seq',
      guide=guide_legend(title='Quality'))
```

Total.sulfer.dioxide appears to have an overall correlation to quality of wine
By rating you can observe the trend going from negative to positive as quality 
goes up. 
summary(redwine)

```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine,
       aes(y = pH, x = alcohol,
           color = quality_rating)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~quality_rating) +
  scale_color_brewer(type='seq',
      guide=guide_legend(title='Quality'))
```
pH and alcohol have a positive correlation at every quality level and a very 
strong correlation at the classic quality level. 

```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine,
       aes(y = sulphates, x = alcohol,
           color = quality_rating)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~quality_rating) +
  scale_color_brewer(type='seq',
      guide=guide_legend(title='Quality'))
```
There does not appear to be a correlation between sulphates, alcohol 
and Quality.

```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine,
       aes(y = sulphates, x = alcohol,
           color = quality_rating)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~quality_rating) +
  scale_color_brewer(type='seq',
      guide=guide_legend(title='Quality'))
```
# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
Sugar and Alcohol were not strengthened. pH and Total.sulfer.Dioxide (TSD) are of 
interest as they show positive correlations. While TSD begins negative you can
observe it becoming more positive through the ratings. 

### Were there any interesting or surprising interactions between features?
I found it interesting that what in some features there were positive 
correlations in lower rated wines but as you skipped to the highest rating 
those would then become positive. This lends me to believe there are 
combinations of features still be be explored. 


------

# Final Plots and Summary
1. Alcohol and pH seems to produce better wines.
2. Residual.Sugar is correlated in wines that are Not recommended, Mediocre and
Classic.

I wanted to stayed with what I knew about wine which was sugar and alcohol. 
While very simple I thought it was a good starting point. 

### Plot One

```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine, aes(x=quality_rating))+
  geom_bar()
```
My initial exploration revealed that the data set is limited to mostly
Good and Very Good Wines. This to me will influence finding correlations with
Quality. 

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library("GGally")
ggcorr(redwine)
```

### Description Two
Luckily I was able
to find a correlation with alcohol but not so much with sugar. 
This led me to review all features.

Alcohol still seems to have the highest relationship on quality, 
followed by volatile.acidity. At this stage it is apparent that other features
will need to be reviewed. 


### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE} 
ggplot(data = redwine,
       aes(y = pH, x = alcohol,
           color = quality_rating)) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
  facet_wrap(~quality_rating) +
  scale_color_brewer(type='seq',
      guide=guide_legend(title='Quality'))
```

### Description Three
While this has a positive Correlation that is not necessarily a good thing. 
After a little review I found that the Higher pH wines will taste flat and lack
freshness. Bacteria also thrives at a higher pH. So why is there a positive
Correlation with higher pH levels and quality of wine. As wine ages the pH level
rises and older wines tend to have higher quality. There is a deeper science
to this and on further research I believe taking a wine data set of 1 variatel
all from the same year would yeild better overal analysis. 
------


## Reflections

This data set was focuse primarly on Good and Very Good wines. With the category
of Quality you are drawn to look for relationships in what would drive quality. 
However when your population is relegated to 2 main categories a lot more 
cleaning would need to be done. 

To build better analysis I would want to normalize each quality level by 
indexing wines based on ratio. 

Upon further reflection I would need to spend more time with my business sponsor
to better understand how pH, Acidity, Chlorides and sulfates work together. 
I believe once I understood this more I would have more confidence in isolating
outliers and removing appropriate ones rather than chopping of a set standard 
quartile. 

I would like to have a dataset that incorporated age, price, variatel. This 
would lend itself to a great classification project. 


